---
title       : Exploratory Analysis of Motor Trend Dataset with Decision Tree 
subtitle    : 
author      : Yuwei Cui
job         : University of Maryland, College Park
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## The Motor Trend Dataset

- The motor trend dataset (mtcars) is a build in dataset in R.
It contains fuel consumption and 10 aspects of automobile design and performance for 32 automobiles.

- Let's load the dataset into R and look at the first three observations:
```{r}
data(mtcars)
head(mtcars,3)
```

- We will build a tool that allows exploratory analysis of the mtcars dataset. You can find the shiny application online [here](http://ywcui.shinyapps.io/MTcar_DecisionTree/).

--- .class #id 

## Decision Tree 

- Decision tree is a decision support tool that uses a tree-like graph to understand consequences of decisions. More information can be found on this [wikipage](http://en.wikipedia.org/wiki/Decision_tree). 
- We used the rpart package to fit decision tree model

```{r}
require(rpart)
modDT <- rpart(mpg ~ wt + hp, data = mtcars)
```

- The formula "mpg ~ wt + hp" informs rpart to use weight (wt) and horsepower (hp) as predictors to explain the difference of fuel consumption (mpg). 

- In the online shiny application, you can specify the predictors and the outcomes from the side panel.

--- .class #id 

## Understanding the Decision Tree

Let's take a look at the decision tree model returned by rpart:
```{r}
print(modDT)
```

The resulting decision tree has 3 levels and 5 nodes. The decision criteria, the number of observations processed by the node, the deviance and the average value of mpg is shown for each node. 

--- .class #id 
## Display Decision Tree

- To visualize the decision tree, we used the fancyRpartPlot function from the rattle package. 
- The same decision tree is plotted in a nice format here. The color of the node indicates the average outcome value (mpg) for that node.
```{r}
fancyRpartPlot(modDT)            
```